Simplifying Lotka--Volterra interaction matrices by species grouping: optimal groupings are unpredictable

### 2022-06-02 article.rnw
Correct the tense of a sentence.
DBS edits and final comments

Dynamic models of ecological communities require many parameters, making them expensive to fit to experimental or observational data.
To reduce the number of parameters, species are often divided into groups \emph{a priori}, using phylogenetic or functional groups, and species within a group are assumed to behave identically.
Here, we assess the validity of such grouping by deriving the optimal groupings \emph{a posteriori}, from the parsimony of a Lotka--Volterra model fit to experimental data from six species of \emph{Drosophila} competing in laboratory conditions.
We find evidence that grouping can improve the parsimony of species interaction models, but we also observe that there is no single optimal grouping.
Further, those groupings that did prove beneficial could not be easily predicted from available species attributes \emph{a priori}, calling into question the prevailing method of \emph{a priori} species grouping.
Lastly, we find that species did not group equivalently in their competitive effect and in their response to competition, invalidating a key assumption often made implicitly when defining species groupings.
This suggests the need for a different, more data-informed, approach to grouping in species interaction models.

% DBS: Once we have a title, make sure that we remove all words from the keywords that appear there.
\keywords{Lotka--Volterra; interaction matrix; competition; grouping; model fitting; parsimony}

@@

Making practical use of a model requires experimentally or observationally determining the values of each of the model's parameters.
This poses a particular problem for models that include species interactions due to the sheer number of parameters that must be determined.
Even the simplest models with no environmental dependence require one parameter to characterise the interaction of each pair of species \parencite{novak2016characterizing,rocha2005lotka}.
% DBS: Hmm. Should the above two or rephrased to embrace the directionality of what's going on?
Hence the number of parameters scales quadratically with the number of species; at least \num{100} parameters are required for the interactions in a community of ``just'' \num{10} species.
Determining these parameters experimentally requires at least one experiment per parameter, and preferably more \parencite{hart2018quantify}.
This is understandably expensive and requires substantial effort.
As such, model-based predictions of real communities are typically restricted to a small number of focal species \parencite{ovaskainen2017species,maynard2020predicting}.
However, real communities can contain hundreds of species \parencite{cornell1999unsaturation}, or even thousands for microbial communities assessed by genetic metabarcoding \parencite{warton2015variables}.
To bridge this gap between the size of real communities and our ability to parameterise models for them, we need an ecologically-defensible way to reduce the number of parameters that predictive models require.

One simplifying approach is to assume that most pairs of species do not interact or interact weakly, and hence setting many of the parameters to zero \parencite{mutshinda2009drives}.
However, weak interactions have been shown to be very important to the stability of ecological communities \parencite{mccann1998weak,neutel2002stability,wootton2016many}, and setting these to zero may lose important information.
Another approach assumes that species fall into groups, such that all species in the same group interact with species in other groups in an identical fashion, thereby collapsing pairwise interaction parameters for all the species in that group into sets of common parameters \parencite{ovaskainen2017species,weiss2021disentangling}.

In ecological contexts, such groups have been defined \emph{a priori} by phylogeny at some taxonomic level or by functional similarity \parencite{uriarte2004spatially,ovaskainen2017species,weiss2021disentangling,martyn2021identifying}.
The argument for grouping based on phylogeny relies on the principle that closely related species will fill similar ecological niches, and hence will interact with other species in a similar fashion \parencite{godoy2014phylogenetic,webb2002phylogenies,darwin1859origin}. %Darwin Chapter 3 "Struggle for Life Most Severe Between Individuals and Varieties of the Same Species".
However, empirical tests of this principle have found remarkably little support for it \parencite{godoy2014phylogenetic,cahill2008does,mayfield2010opposing}.
On the other hand, functional groupings have also proven to be poor predictors of species interactions \parencite{bret2008plant,funk2017revisiting}.
While attempts have been made to define functional groupings that are more useful in a modelling context \parencite{boulangeat2012improving}, these rely upon measurements of biotic interactions to define such groups.
As such, integrating these into species interaction models leaves us with a circular argument: a functional group is a group of species that interact with others in a similar way, and we therefore assume that all the species in a functional group interact with others in a similar way.
In all, we are left with the question of how best to group species in community models, and whether we can do so in a meaningful way using less data than required for a fully pairwise parameterisation.

To the best of our knowledge, all studies that use grouping to simplify models of species interactions have also made one additional, untested assumption: that the same grouping is equally valid for how species affect others as for how species respond to others.
There is good reason to question this assumption, as it is known that different traits can govern species' competitive effects than govern their responses to competition \parencite{goldberg1991competitive}.
For resource competition in particular, competitive effect is determined by how the organism affects local resource availability, while competitive response is determined by the organism is inhibited by depleted resources \parencite{violle2009competition}.
As such, the best groupings of species may be different for each of the effects and responses.

We propose that species might be best partitioned into groups without regard to any \emph{a priori} definition but instead based purely upon which grouping gives the most parsimonious model.
Data-informed grouping approaches have previously proved successful in other contexts, such as finding modules in networks \parencite{guimera2005functional,newman2006modularity,girvan2002community}.
In this study, we investigate the usefulness and correctness of such an \emph{a posteriori} approach using experimental data from competition between different \emph{Drosophila} species provided by \textcite{terry2021natural}.
We fit a Lotka--Volterra model to these data to find the groupings that give the most parsimonious model, and compare these groupings to similarity based upon a species trait (body size).
While doing so, we group species independently with regard to their effects upon other species and their responses to other species, to find whether the data-informed groupings differ between these two cases.

@@

The data we used in this study were drawn from the work of \textcite{terry2021natural}.
The data consist of fecundity measurements---number of surviving offspring---for six species of \emph{Drosophila}, grown in single-species or two-species communities with varying densities of each species.
This experimental design corresponds to a response--surface experimental design commonly used to parameterise a species interaction model \parencite{hart2018quantify,zhang2019common,cameron2019size}.

@@

@@

The data provided the change in population size from one generation to the next, so we used a discrete-time version of the Lotka--Volterra competition model given by:
\begin{equation}
	N_{i,t+1} - N_{i,t} = N_{i,t} \, r_i \left(1 - \sum_j \alpha_{ij} N_{j,t}\right) \,,
	\label{eq:lv-model}
\end{equation}
where $N_{i,t}$ is the population of species $i$ at time $t$ (the size of the parent generation), $N_{i,t+1}$ is the population at time $t+1$ (the size of the offspring generation), $r_i$ is the intrinsic growth rate of species $i$, and $\alpha_{ij}$ is the per capita effect of species $j$ on species $i$.
The competition coefficients were not constrained to be positive, allowing for the possibility of mutualistic effects between some species.
With $n$ species, this model has $n+n^2$ parameters: $n$ growth rates and $n^2$ competition coefficients.

@@

In order to reduce the number of free parameters in the model, we grouped species together, and constrained the appropriate competition coefficients to be equal for species within the same group.
Within the model defined above, each species is effectively in its own group, such that all pairs of species interact in a unique fashion.
At the other extreme, all species could be placed in the same group, such that all pairs of species interact equivalently; they are ecologically equivalent \emph{sensu} \textcite{mcpeek2019disentangling}.
Given $n$ species, there may be any integer number of groups from $1$ to $n$.
Groupings are known formally as partitions in mathematics, and the total number of groupings for $n$ species is given by the Bell numbers, denoted $B_n$ \parencite{bell1934exponential}.

It is commonly assumed that species which are similar in their responses to competition would automatically be similar in their competition effects.
However, competitive responses may be governed by a different set of traits than competitive effects, so this assumption does not necessarily hold.
We avoided this assumption by allowing species which were similar in one regard to be dissimilar in the other, and vice versa.
As such, each model was defined by two groupings of species: one grouping of the competitive responses, and one of the competitive effects.

To better understand the meaning of species grouping of competition coefficients, it is useful to recognise that this component of the Lotka--Volterra model can be represented as an interaction matrix \parencite{novak2016characterizing}, where a row contains all the competitive effects on a single species and a column contains all the competitive effects of a single species.
As such, if two species are assumed to have the same response to competition, then two rows of the interaction matrix are constrained to be equal.
Similarly, if two species are assumed to have the same competitive effect, then two columns of the interaction matrix are constrained to be equal.
This allows us to refer below to the grouping of the competitive responses as the ``row grouping'' and the grouping of the competitive effects as the ``column grouping''.
Formally, if species $a$ and $b$ were in the same row group, $\alpha_{aj}\equiv\alpha_{bj} \, \forall j$, and if $a$ and $b$ were in the same column group, $\alpha_{ia}\equiv\alpha_{ib} \, \forall i$.

@@

Using these Akaike weights, we calculated two important statistics.
First, the total Akaike weight of models with a lower {\ic} than the entirely ungrouped model; this is the probability that some model is more parsimonious than the fully ungrouped model.
Second, the total Akaike weight of models that used the same grouping for the rows as for the columns; this is the probability that the most parsimonious model grouped species equivalently in their competitive effects as in their competitive responses.
To evaluate the significance of the latter, we performed a Monte Carlo permutation test, comparing the total Akaike weight of those models that grouped rows and columns equivalently to the total Akaike weight of a equally sized random subset of the models.

@@

However, we could test the row-grouping and column-grouping coclassification matrices against one another.
By testing for correlation between the coclassification matrix for the the rows and that for the columns, we could perform a different test regarding the assumption that it is appropriate to use the same grouping for both.
We did this using a Mantel test, as appropriate for similarity matrices.
We could also test the coclassification matrices for correlation with an expected dissimilarity matrix.
We calculated one such dissimilarity matrix based on a species trait: body size (as mean wet mass of female adults, obtained from \cite{terry2021natural}).
The trait difference matrix was calculated as absolute difference between trait values, while the coclassification matrices were converted to dissimilarity matrices by subtracting them from $1$.
We then performed Mantel tests for correlation between the trait distance matrix and each of the two coclassification dissimilarity matrices.

@@

% DBS: Semi-random thought: should we also add the "all alphas = 0" model to the mix?

@@

\begin{figure}
	<<>>=
		hist(brute_data$statistics[[ic]],
			main="",
			xlab=ic_name,
			xaxs='i',
			yaxs='i'
		)
		abline(v = brute_data$fully_separated_statistics[[ic]], lwd=2, lty="dashed", col="green")
		abline(v = brute_data$fully_grouped_statistics[[ic]], lwd=2, lty="dashed", col="red")
	@

	\caption{%
		Histogram of the {\ic} values of Lotka--Volterra model fit with all possible groupings of interaction coefficients.
		The green line shows the {\ic} value (\icnum{\Sexpr{brute_data$fully_separated_statistics[[ic]]}}) of the ungrouped model, in which all competition coefficients are fitted independently.
		The red line shows the {\ic} value (\icnum{\Sexpr{brute_data$fully_grouped_statistics[[ic]]}}) of the fully grouped model, where every pair of species shares a single interaction coefficient.
	}
	\label{fig:ic}
\end{figure}

@@

Considering only models in which the rows and columns were grouped using the same grouping, the {\ic} values ranged from \icnum{\Sexpr{min(brute_data$equivalently_grouped_statistics[[ic]])}} to \icnum{\Sexpr{max(brute_data$equivalently_grouped_statistics[[ic]])}}; neither the best nor worst overall models had equivalently grouped rows and columns.
The total Akaike weight assigned to these models was \probnum{\Sexpr{brute_data$equivalently_grouped_weight}}; this was more than expected by chance (\probnum{\Sexpr{brute_data$equivalently_grouped_expected_weight}}), but not significantly so (Monte Carlo permutation test, $p=\probnum{\Sexpr{brute_data$equivalently_grouped_weight_p #$}}$).
This indicates that the most parsimonious model is not especially likely to be one that groups the rows and columns equivalently.

@@

	\begin{subfigure}{\textwidth}
		\centering
		<<>>=
			grouped_alpha_matrix(
				best_species_names,
				best_row_grouping,
				best_col_grouping,
				best_alpha_matrix
			)
		@
		\caption{The grouped alpha matrix.}
		\label{fig:best-alpha-matrix}
	\end{subfigure}

@@

	\begin{subfigure}{0.49\textwidth}
		\centering
		<<>>=
			minimal_coclassification_kable(
				best_col_coclassification,
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{The coclassification matrix of rows.}
		\label{fig:best-coclassification-cols}
	\end{subfigure}
	%
	\begin{subfigure}{0.49\textwidth}
		\centering
		<<>>=
			minimal_coclassification_kable(
				best_row_coclassification,
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{The coclassification matrix of columns.}
		\label{fig:best-coclassification-rows}
	\end{subfigure}

	\caption{%
		The groupings and competition coefficients of a grouped Lotka--Volterra model fitted to the six studied \emph{Drosophila} species.
		This represents the single model with the best {\ic} (\probnum{\Sexpr{brute_data$statistics[best_model_index,ic]}}).
		\Subref{fig:best-alpha-matrix} shows the competition coefficients, where a competition coefficient in a block represents the same coefficient being used for multiple interactions, per the groupings.
		\Subref{fig:best-coclassification-cols} and \subref{fig:best-coclassification-rows} show the coclassification matrices derived from the groupings of the columns and rows respectively.
		These coclassification matrices are averaged for all models, weighted by Akaike weights, to produce \cref{fig:coclassification}.
		Species have been reordered that species which are grouped together appear adjacent.
		The ordering is the same in \subref{fig:best-coclassification-cols} and \subref{fig:best-coclassification-rows} as shown in \subref{fig:best-alpha-matrix}.
	}
	\label{fig:best-model}
\end{figure}

@@

Overall, no pair of species can confidently be said to be coclassified together in the row or column groupings; the highest probability of classification was \SI[round-mode=figures,round-precision=3]{\Sexpr{max(weighted_col_coclassification_matrix - diag(num_species), weighted_row_coclassification_matrix - diag(num_species))*100}}{\percent} (\cref{fig:coclassification}).
However, some pairs of species were far more likely to be coclassified than others, suggesting some structure to the best groupings.

The structure of the coclassifications was very different in the row groups than in the column groups.
Most pairs of species which had a high probability of being coclassified in one grouping had a very low probability of being coclassified in the other grouping.
Only one pair of species had a relatively high probability of being coclassified in both groupings: \emph{D. pallidifrons} and \emph{D. sulfurigaster}.
Overall, more coclassification occurred in the row groupings than in the column groupings.
A Mantel test of the coclassification matrix for the rows against that for the columns shows that there is no significant correlation between the two ($\mantelstat=\cornum{\Sexpr{mantel_cor(weighted_col_coclassification_matrix, weighted_row_coclassification_matrix)}}$, $p=\probnum{\Sexpr{mantel_p(weighted_col_coclassification_matrix, weighted_row_coclassification_matrix)}}$).

@@

% DBS: Most journals frown upon or do not allow overly verbose subtitles on figures. So I've cut them down to be a bit more concise.
% DBS: But the fact that the labels say effects and responses reminds me of a lingering "doubt" I had when reading the main text. We talk almost entirely about rows and columns but there are a number of junctures where I feel like it would be easier to follow for people who aren't us if we were talking about responses and effects. As long as we're consistent, I'm fine either way. But we need to remember we're writing the paper for the audience and not for ourselves. ;)
\begin{figure}
	\centering
	\begin{subfigure}{\textwidth}
		\centering
		<<>>=
			weighted_coclassification_kable(
				weighted_col_coclassification_matrix[species_output_order,species_output_order],
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{Coclassification of competitive effects}
		\label{fig:coclassification-cols}
	\end{subfigure}

@@

	\caption{%
		Coclassification matrices of the Lotka--Volterra competitive effects and responses of the six studied \emph{Drosophila} species, weighted by Akaike weights.
		The cell for a pair of species represents the probability that the best grouping, as determined by the Akaike weights of the grouped models, places those two species in the same group.
		\Subref{fig:coclassification-cols} represents the grouping of species' competitive effects upon other species (columns of the interaction matrix), while \subref{fig:coclassification-rows} represents the grouping of species' responses to competition (rows of the interaction matrix).
	}
	\label{fig:coclassification}
\end{figure}

% DBS: I think a good figure would be to show the ungrouped estimated alpha matrix in one panel and the alpha matrix averaged across all partitions in a second panel, with lines drawn in the second to reflect the "optimal" partitioning of the co-classification matrices.

@@

We have shown that it is possible to fit a more parsimonious model of species interactions by grouping the interaction coefficients of species \emph{a posteriori}, and hence reducing the number of free parameters in the model.
Indeed, the results show that grouping species was likely to give a more parsimonious model (probability \probnum{\Sexpr{sum(brute_data$statistics[[ic_weight]][brute_data$statistics[[ic]] < brute_data$fully_separated_statistics[[ic]]])}}) than leaving them ungrouped.
Which grouping would give the most parsimonious model, however, could not be determined with confidence.
This result stands in contrast to the popular wisdom in ecology that species should form some sort of clear groups, be they the functional groups so often used \parencite{blondel2003guilds,kirwan2009diversity,boulangeat2012improving}, or taxonomic groups per a relationship between phylogeny and ecological niche \parencite{godoy2014phylogenetic,wilcox2018evolutionary}.
It also contrasts with previous studies that have succeeded in finding clear groups in other ecological contexts, such as in food webs \parencite{krause2003compartments,allesina2009food} and pollination networks \parencite{olesen2007modularity}.

>>>>

Most dynamic models of ecological communities require many parameters, making them expensive to fit to experimental or observational data.
To reduce the number of parameters, species are often divided into groups \emph{a priori}, using phylogenetic or functional groups, and species within a group are assumed to behave identically.
Here, we assess the validity of such grouping by deriving the optimal groupings \emph{a posteriori} based on the fit of the Lotka--Volterra competition model to experimental data from six species of \emph{Drosophila} competing in laboratory conditions.
We find evidence that grouping can improve the parsimony of species interaction models, but we also observe that there is no single optimal grouping.
Further, those groupings that did prove beneficial could not have been easily predicted from available species attributes, calling into question the prevailing method of \emph{a priori} species grouping.
Lastly, we find that species did not group equivalently in their competitive effect and in their response to competition, invalidating a key assumption often made implicitly when defining species groupings.
Our study suggests the need for a different, more data-informed, approach to grouping in species interaction models.

% DBS: Once we have a title, make sure that we remove all words from the keywords that appear there.
\keywords{competition; ecological equivalence; grouping; model fitting; parsimony; \emph{Drosophila}}

@@

Making practical use of a model requires experimentally or observationally determining the values of each of the model's parameters.
This poses a particular problem for models that include species interactions due to the sheer number of parameters that must be determined.
Even the simplest models with no environmental dependence require one parameter to characterise the interaction of each pair of species \parencite{novak2016characterizing,rocha2005lotka}.
Hence the number of parameters scales quadratically with the number of species; at least \num{100} parameters are required for the interactions in a community of ``just'' \num{10} species.
Determining these parameters experimentally requires at least one experiment per parameter, and preferably more \parencite{hart2018quantify}.
This is understandably expensive and requires substantial effort.
As such, model-based predictions of real communities are typically restricted to a small number of focal species \parencite{ovaskainen2017species,maynard2020predicting}.
However, real communities can contain hundreds of species \parencite{cornell1999unsaturation}, or even thousands for microbial communities assessed by genetic metabarcoding \parencite{warton2015variables}.
To bridge this gap between the size of real communities and our ability to parameterise models for them, we need an ecologically-defensible way to reduce the number of parameters that predictive models require.

One simplifying approach is to assume that most pairs of species do not interact or interact weakly, and hence setting many of the parameters to zero \parencite{mutshinda2009drives}.
However, weak interactions have been shown to be very important to the stability of ecological communities \parencite{mccann1998weak,neutel2002stability,wootton2016many}, and setting any individual parameter to zero risks losing important information.
Another approach assumes that species fall into groups, such that all species in the same group interact with species in other groups in an identical fashion, thereby collapsing pairwise interaction parameters for all the species in that group into sets of common parameters \parencite{ovaskainen2017species,weiss2021disentangling}.

In ecological contexts, such groups have been defined \emph{a priori} by phylogeny at some taxonomic level or by functional (i.e., trait) similarity \parencite{uriarte2004spatially,ovaskainen2017species,weiss2021disentangling,martyn2021identifying}.
The argument for grouping based on phylogeny relies on the principle that closely related species will fill similar ecological niches, and hence will interact with other species in a similar fashion \parencite{godoy2014phylogenetic,webb2002phylogenies,darwin1859origin}. %Darwin Chapter 3 "Struggle for Life Most Severe Between Individuals and Varieties of the Same Species".
However, empirical tests of this principle have found remarkably little support for it \parencite{godoy2014phylogenetic,cahill2008does,mayfield2010opposing}.
On the other hand, functional groupings have also proven to be poor predictors of species interactions \parencite{bret2008plant,funk2017revisiting}.
While attempts have been made to define functional groupings that are more useful in a modelling context \parencite{boulangeat2012improving}, these rely upon measurements of biotic interactions to define such groups.
As such, integrating these into species interaction models leaves us with a circular argument: a functional group is a group of species that interact with others in a similar way, and we therefore assume that all the species in a functional group interact with others in a similar way.
In all, we are left with the question of how best to group species in community models, and whether we can do so in a meaningful way using less data than required for a fully pairwise parameterisation.

To the best of our knowledge, all studies that use grouping to simplify models of species interactions have also made one additional, untested assumption: that the same grouping is equally valid for how species affect others as for how species respond to others.
There is good reason to question this assumption, as it is known that different traits can govern species' competitive effects than govern their responses to competition \parencite{goldberg1991competitive}.
For resource competition in particular, competitive effects are determined by how the organism affects local resource availability while competitive responses are determined by how the organism is inhibited by depleted resources \parencite{violle2009competition}.
As such, the best groupings of species may be different for each of the effects and responses.

We propose that species might be best partitioned into groups without regard to any \emph{a priori} definition but instead based purely upon which grouping gives the most parsimonious model.
Such data-informed grouping approaches have previously proved successful in other contexts, such as finding modules in networks \parencite{guimera2005functional,newman2006modularity,girvan2002community}.
In this study, we investigate the usefulness and correctness of such an \emph{a posteriori} approach using experimental data from competition between different \emph{Drosophila} species provided by \textcite{terry2021natural}.
We fit a Lotka--Volterra competition model to these data to find the groupings that give the most parsimonious model, and we then compare these groupings to similarity based upon a species trait (body size).
While doing so, we group species independently with regard to their effects upon other species and their responses to other species to find whether the data-informed groupings differ between these two cases.

@@

The data used in this study were drawn from the work of \textcite{terry2021natural}.
They consist of fecundity measurements---number of surviving offspring---for six species of \emph{Drosophila}, grown in single-species or two-species communities with varying densities of each species.
This experimental design corresponds to a response--surface experimental design commonly used to parameterise species interactions in population dynamics models \parencite{hart2018quantify,zhang2019common,cameron2019size}.

@@

% DBS: Wait, the response is pairs? So can we have non-integer responses?

@@

The data provided the change in population size from one generation to the next, so we used a discrete-time version of the Lotka--Volterra competition model given by:
\begin{equation}
	N_{i,t+1} - N_{i,t} = N_{i,t} \, r_i \left(1 - \sum_j^n \alpha_{ij} N_{j,t}\right) \,,
	\label{eq:lv-model}
\end{equation}
where $N_{i,t}$ is the population of species $i$ at time $t$ (the size of the parent generation), $N_{i,t+1}$ is the population at time $t+1$ (the size of the offspring generation), $r_i$ is the intrinsic growth rate of species $i$, and $\alpha_{ij}$ is the per capita effect of species $j$ on species $i$.
When fitting the data to the model, the competition coefficients were not constrained to be positive; this allows for the possibility of mutualistic effects between some species.
With $n$ species, this model has $n+n^2$ parameters: $n$ growth rates and $n^2$ competition coefficients.

% DBS: A tardy thought, if we don't constrain alphas I believe many people would say we should claim that we are fitting the generalised Lotka--Volterra model and *not* a Lotka--Volterra competition model.

@@

In order to reduce the number of free parameters in the model, we grouped species together, and constrained the appropriate competition coefficients to be equal for species within the same group.
Within the model defined above, each species is effectively in its own group, such that all pairs of species interact in a unique fashion.
At the other extreme, all species could be placed in the same group, such that all pairs of species interact equivalently; they are ecologically equivalent \emph{sensu} \textcite{mcpeek2019disentangling}.
Given $n$ species, there may be any integer number of groups from $1$ to $n$.
Groupings are known formally as partitions in mathematics, and the total number of ways to group $n$ species is given by the Bell number, denoted $B_n$ \parencite{bell1934exponential}.

It is commonly assumed that species which are similar in their responses to competition would automatically be similar in their competition effects.
However, competitive responses may be governed by a different set of traits than competitive effects, so this assumption may not always hold.
We therefore relaxed this assumption by allowing species which were similar in one regard to be dissimilar in the other, and vice versa.
As such, each grouped model was defined by two groupings of species: one grouping of the competitive responses and one of the competitive effects.

To better understand the meaning of species grouping of competition coefficients, it is useful to recognise that this component of the Lotka--Volterra competition model can be represented as an interaction matrix \parencite{novak2016characterizing}, where a row contains all the competitive effects on a single species and a column contains all the competitive effects of a single species.
If two species are assumed to have the same responses to competition, then two rows of the interaction matrix are constrained to be equal.
Similarly, if two species are assumed to have the same competitive effects, then two columns of the interaction matrix are constrained to be equal.
This allows us to refer below to the grouping of the competitive responses as the ``row grouping'' and the grouping of the competitive effects as the ``column grouping''.
Formally, if species $a$ and $b$ were in the same row group, $\alpha_{aj}\equiv\alpha_{bj} \, \forall j$, and if $a$ and $b$ were in the same column group, $\alpha_{ia}\equiv\alpha_{ib} \, \forall i$.

@@

Using these Akaike weights, we calculated two important statistics.
First, the total Akaike weight of models with a lower {\ic} than the entirely ungrouped model; this is the probability that some model is more parsimonious than the fully ungrouped model.
Second, the total Akaike weight of models that used the same grouping for the rows as for the columns; this is the probability that the models grouping species equivalently in their competitive effects and competitive responses were more parsimonious than all other models.
To evaluate the significance of the latter, we performed a Monte Carlo permutation test, comparing the total Akaike weight of those models that grouped rows and columns equivalently to the total Akaike weight of a equally sized random subset of the models.

@@

However, we could test the row-grouping and column-grouping coclassification matrices against one another.
By testing for correlation between the coclassification matrix for the the rows and that for the columns, we could perform a second test of the assumption that it is appropriate to use the same grouping for both.
We did this using a Mantel test, as appropriate for similarity matrices.
We could also test the coclassification matrices for correlation with an expected dissimilarity matrix.
We calculated one such dissimilarity matrix based on a species trait: body size (as mean wet mass of female adults, obtained from \cite{terry2021natural}).
The trait difference matrix was calculated as absolute difference between trait values while the coclassification matrices were converted to dissimilarity matrices by subtracting them from $1$.
We then performed Mantel tests for correlation between the trait distance matrix and each of the row-grouping and column-grouping coclassification dissimilarity matrices.

@@

% DBS: We need to add a reference to the partial residual plots in the supps.

@@

\begin{figure}
	\centering
	\begin{subfigure}{0.65\textwidth}
	<<>>=
		hist(brute_data$statistics[[ic]],
			main="",
			xlab=ic_name,
			xaxs='i',
			yaxs='i',
			ylim=c(0,6000)
		)
		abline(v = brute_data$fully_separated_statistics[[ic]], lwd=2, lty="dotted", col="green")
		abline(v = brute_data$fully_grouped_statistics[[ic]], lwd=2, lty="dashed", col="red")
	@
	\end{subfigure}

	\caption{%
		Histogram of the {\ic} values of Lotka--Volterra model fits with all possible groupings of interaction coefficients.
		The dotted green line shows the {\ic} value (\icnum{\Sexpr{brute_data$fully_separated_statistics[[ic]]}}) of the ungrouped model in which all competition coefficients are fitted independently.
		The dashed red line shows the {\ic} value (\icnum{\Sexpr{brute_data$fully_grouped_statistics[[ic]]}}) of the fully grouped model in which every pair of species shares a single interaction coefficient.
	}
	\label{fig:ic}
\end{figure}
% DBS: Reconsider green and red to be more colour-blind friendly. I've already changed it so that the lines are different types.

@@

Considering only models in which the rows and columns were grouped using the same grouping, the {\ic} values ranged from \icnum{\Sexpr{min(brute_data$equivalently_grouped_statistics[[ic]])}} to \icnum{\Sexpr{max(brute_data$equivalently_grouped_statistics[[ic]])}}; neither the best nor worst overall models had equivalently grouped rows and columns.
The total Akaike weight assigned to these models was \probnum{\Sexpr{brute_data$equivalently_grouped_weight}}; this was more than expected by chance (\probnum{\Sexpr{brute_data$equivalently_grouped_expected_weight}}), but not significantly so (Monte Carlo permutation test, $p=\probnum{\Sexpr{brute_data$equivalently_grouped_weight_p #$}}$).
This indicates that the most parsimonious model was not especially likely to be one that grouped the rows and columns equivalently.

@@

	\begin{subfigure}{\textwidth}
		\centering
		<<>>=
			grouped_alpha_matrix(
				best_species_names,
				best_row_grouping,
				best_col_grouping,
				best_alpha_matrix
			)
		@
		\caption{Grouped alpha matrix}
		\label{fig:best-alpha-matrix}
	\end{subfigure}

@@

	\begin{subfigure}{0.49\textwidth}
		\centering
		<<>>=
			minimal_coclassification_kable(
				best_col_coclassification,
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{Coclassification matrix of rows}
		\label{fig:best-coclassification-cols}
	\end{subfigure}
	%
	\begin{subfigure}{0.49\textwidth}
		\centering
		<<>>=
			minimal_coclassification_kable(
				best_row_coclassification,
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{Coclassification matrix of columns}
		\label{fig:best-coclassification-rows}
	\end{subfigure}

	\caption{%
		The groupings and competition coefficients of a grouped Lotka--Volterra model fitted to the six studied \emph{Drosophila} species.
		Where results are for the single model with the best {\ic} (\probnum{\Sexpr{brute_data$statistics[best_model_index,ic]}}).
		\Subref{fig:best-alpha-matrix} We show the inferred matrix of competition coefficients, where a competition coefficient in a block represents the same coefficient being used for multiple interactions (per the identified groupings).
		\Subref{fig:best-coclassification-cols} and \subref{fig:best-coclassification-rows} We show the coclassification matrices derived from the groupings of the columns and rows, respectively.
		%These coclassification matrices are averaged for all models, weighted by Akaike weights, to produce \cref{fig:coclassification}.
		Species have been reordered so that species that were grouped together appear adjacent, and the ordering is the same in all of \subref{fig:best-alpha-matrix}, \subref{fig:best-coclassification-cols}, and \subref{fig:best-coclassification-rows}.
	}
	\label{fig:best-model}
\end{figure}

% DBS: I cut out a sentence above in the caption because we say this in the text elsewhere and it's a bit unusual to refer to a later figure in the caption of an earlier one.

@@

Overall, no pair of species could confidently be said to be coclassified together in the row or column groupings; the highest probability of classification was \SI[round-mode=figures,round-precision=3]{\Sexpr{max(weighted_col_coclassification_matrix - diag(num_species), weighted_row_coclassification_matrix - diag(num_species))*100}}{\percent} (\cref{fig:coclassification}).
However, some pairs of species were far more likely to be coclassified than others, suggesting some structure to the best groupings.

The structure of the coclassification matrices was very different in the row groups than in the column groups.
That is, most pairs of species which had a high probability of being coclassified in one grouping had a very low probability of being coclassified in the other grouping.
Only one pair of species had a relatively high probability of being coclassified in both groupings: \emph{D. pallidifrons} and \emph{D. sulfurigaster}.
Overall, more coclassification occurred in the row groupings than in the column groupings.
A Mantel test of the coclassification matrix for the rows against that for the columns showed that there is no significant correlation between the two ($\mantelstat=\cornum{\Sexpr{mantel_cor(weighted_col_coclassification_matrix, weighted_row_coclassification_matrix)}}$, $p=\probnum{\Sexpr{mantel_p(weighted_col_coclassification_matrix, weighted_row_coclassification_matrix)}}$).

@@

% DBS: But the fact that the labels say effects and responses reminds me of a lingering "doubt" I had when reading the main text. We talk almost entirely about rows and columns but there are a number of junctures where I feel like it would be easier to follow for people who aren't us if we were talking about responses and effects. As long as we're consistent, I'm fine either way. But we need to remember we're writing the paper for the audience and not for ourselves. ;)

% DBS: Adding a new note here. Shouldn't we pick one version that gets used in Fig 2 and in Fig 3 to refer to the two coclassification matrices?

\begin{figure}
	\centering
	\begin{subfigure}{\textwidth}
		\centering
		<<>>=
			weighted_coclassification_kable(
				weighted_col_coclassification_matrix[species_output_order,species_output_order],
				colourmap="coclassification", diagonal_colour="coclassification diagonal"
			)
		@
		\caption{Coclassification of competitive effects}
		\label{fig:coclassification-cols}
	\end{subfigure}

@@

	\caption{%
		Coclassification matrices of the Lotka--Volterra competitive effects and responses of the six studied \emph{Drosophila} species, weighted by Akaike weights.
		The cell for a pair of species represents the probability that the best grouping, as determined by the Akaike weights of the grouped models, places those two species in the same group.
		\Subref{fig:coclassification-cols} We show the grouping of species' competitive effects upon other species (columns of the interaction matrix). \subref{fig:coclassification-rows} We show the grouping of species' responses to competition (rows of the interaction matrix).
	}
	\label{fig:coclassification}
\end{figure}

@@

We have shown that it is possible to fit a more parsimonious model of species interactions by grouping the interaction coefficients of species \emph{a posteriori}, and hence reducing the number of free parameters in the model.
Indeed, the results show that grouping species was likely to give a more parsimonious model (probability \probnum{\Sexpr{sum(brute_data$statistics[[ic_weight]][brute_data$statistics[[ic]] < brute_data$fully_separated_statistics[[ic]]])}}) than was leaving them ungrouped.
Which grouping would give the most parsimonious model, however, could not be determined with confidence.
This result stands in contrast to the popular wisdom in ecology that species should form some sort of clear groups, be they the functional groups so often used \parencite{blondel2003guilds,kirwan2009diversity,boulangeat2012improving}, or taxonomic groups per a relationship between phylogeny and ecological niche \parencite{godoy2014phylogenetic,wilcox2018evolutionary}.
It also contrasts with previous studies that have succeeded in finding clear groups in other ecological contexts, such as in food webs \parencite{krause2003compartments,allesina2009food} or plant--pollinator networks \parencite{olesen2007modularity}.

One possible explanation for our results is that species simply do not group to the same extent in a quantitative competitive context as in a binary interaction networks.
That said, the results also suggest that this is unlikely as grouping was useful for increasing parsimony.
In the present case, we cannot rule out the possibility that a lack of clarity in the groupings is instead a product of the study system.
All the species in this study were members of the same genus, \emph{Drosophila}, and as such were all rather similar.
A study system with species spanning multiple functional groups, or a broader taxonomy, might have shown less equivocal groupings.
Alternatively, the small number of species in the study might not have been sufficient for clear groups to be defined.
Alas, the brute force method we employed in this study has a very poor algorithmic complexity, preventing an application of this approach to studies with more than seven or eight species (there are ${\left(B_7\right)}^2 = \num{\Sexpr{numbers::bell(7)^2}}$ possible models with $7$ species and ${\left(B_8\right)}^2 = \num{\Sexpr{numbers::bell(8)^2}}$ models with $8$ species).
As such, a better method would be required in order to test for grouping of species interactions among a larger set of species.
